\begin{abstract}
\label{abstract}

The multi-axis force sensor market is currently dominated by strain gauge sensors, but other options exist, such as hall-effect, fiber Bragg grating (FBG), and optoelectronic cells. 


The optoelectronic cell is perceptible due to its low cost, excellent precision, and simple implementation. Nevertheless, a dearth of solutions appropriate for large-scale manufacturing and enable adjustable sensor parameters exists.


This study aims to explore how the shape and material of the barrier affect the measurement range of an optoelectronic barrier-type sensor. Additionally, a modular multi-axis force sensor will be designed using the measurement cells.


The study design involves examining the impact of barrier shape and material on the measurement range of the optoelectronic barrier-type sensor. Additionally, in the research, I designed a modular multi-axis force sensor based on the presented measurement cells.


The optoelectronic six-axial force sensor, which is the focus of this study, is characterized by its simplicity and ease of implementation as a modular sensor. It provides a cost-effective solution and facilitates the straightforward replacement of construction elements.
 Consequently, this sensor has the potential to broaden the range of applications for multi-axis force sensors.

This research contributes to understanding how the shape and material of the barrier affect the measurement range of optoelectronic barrier-type sensors. The design of a modular multi-axis force sensor provides a cost-effective and adjustable solution for various robotic applications.
\end{abstract}